A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-04 paper.Rmd
small typos in section 5.6

In regression predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable Y given the value of the independent variable X, $E(Y|X)$. This is an often forgotten element of regression analysis but it is important. Where X is observed, particularly the distribution of the X values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots. This experiment was constructed to assess this, based on four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed). We would expect that if $\hat{Y}$ has a uniform distribution, this would make it easier to read the relationship with the residuals.

@@

```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests. Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations. The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines. The fitted value distribution has greater impact on the power of conventional tests than visual tests. Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450, fig.pos="t"}

>>>>

In regression analysis, predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable $Y$ given the value of the independent variable $X$, $E(Y|X)$. This is an often forgotten element of regression analysis but it is important. Where $X$ is observed, particularly the distribution of the $X$ values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots. This experiment was constructed to assess this, based on four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed). We would expect that if $\hat{Y}$ has a uniform distribution, this would make it easier to read the relationship with the residuals.

@@

```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns. The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests, respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey). Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests. Uneven distributions, normal and lognormal distributions, tend to yield lower power.", dev = 'png', dpi = 450, fig.pos="t"}

### 2023-04-04 paper.tex
small typos in section 5.6

In regression predictions are conditional on the observed values of the
predictors, that is, the conditional mean of the dependent variable Y
given the value of the independent variable X, \(E(Y|X)\). This is an
often forgotten element of regression analysis but it is important.
Where X is observed, particularly the distribution of the X values in
the sample, or consequently \(\hat{Y}\), may affect the ability to read
any patterns in the residual plots. This experiment was constructed to
assess this, based on four different distributions of fitted values:
uniform, normal, discrete and lognormal (skewed). We would expect that
if \(\hat{Y}\) has a uniform distribution, this would make it easier to
read the relationship with the residuals.

>>>>

In regression analysis, predictions are conditional on the observed
values of the predictors, that is, the conditional mean of the dependent
variable \(Y\) given the value of the independent variable \(X\),
\(E(Y|X)\). This is an often forgotten element of regression analysis
but it is important. Where \(X\) is observed, particularly the
distribution of the \(X\) values in the sample, or consequently
\(\hat{Y}\), may affect the ability to read any patterns in the residual
plots. This experiment was constructed to assess this, based on four
different distributions of fitted values: uniform, normal, discrete and
lognormal (skewed). We would expect that if \(\hat{Y}\) has a uniform
distribution, this would make it easier to read the relationship with
the residuals.
